---
title: "ai_model_assignment"
author: "Katie Willi and Caitlin Mothes"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

For this assignment, copy and paste both **your question to the chat bot** as well as **the chat bot's response** below each assignment question.

1)  Ask an AI chat bot (like ChatGPT or Claude) what the following chunk of code does:

```{r}
library(palmerpenguins)

data("penguins")

step_one <- function(x){
  x <- as.character(x)
  }

species_list <- unlist(lapply(penguins$species, step_one))
adelie_indices <- which(species_list == "Adelie")
adelie_penguins <- penguins[adelie_indices, ]
```

**Question posed to chat bot:**

"what does this chunk of code do?"

**Chat bot response:**

This R code performs the following steps:

1\. \*\*Loading [[Data:\\\\\\\*\\\\](Data:\\\*\\){.uri}]([Data:\\\*\\](Data:\*\){.uri}){.uri}\*

```{r}
   library(palmerpenguins)

   data("penguins")
```

It loads the \`palmerpenguins\` package and imports the \`penguins\` dataset into the R environment.

2\. \*\*Defining a Function:\*\*

```{r}
   step_one <- function(x){

     x <- as.character(x)

   }
```

It defines a function named \`step_one\`. However, this function doesn't have a return statement, so it doesn't modify or return any values.

3\. \*\*Applying the Function and Filtering [[Data:\\\\\\\*\\\\](Data:\\\*\\){.uri}]([Data:\\\*\\](Data:\*\){.uri}){.uri}\*

```{r}
species_list <- unlist(lapply(penguins$species, step_one))
```

It applies the \`step_one\` function to convert the \`species\` column of the \`penguins\` dataset into a character vector and stores the result in \`species_list\`.

4\. \*\*Finding Adelie Penguins:\*\*

```{r}
   adelie_indices <- which(species_list == "Adelie")

   adelie_penguins <- penguins[adelie_indices, ]
```

It finds the indices in \`species_list\` where the species is "Adelie". Then, it uses these indices to subset the original \`penguins\` dataset and stores the resulting subset in \`adelie_penguins\`. \`adelie_penguins\` now contains only the rows where the species is "Adelie".

In summary, this code snippet filters the \`penguins\` dataset to include only the rows where the species is "Adelie" and stores this subset in the variable \`adelie_penguins\`. The intermediate steps involve converting the species column to character type, although this doesn't seem to have an impact on the final result since the function \`step_one\` doesn't modify the data.

</br>

2)  Ask the AI model to simplify the code in question 1.

**Question to chat bot:**

"can you simplify the code above?"

**Chat bot response:**

Certainly! The given code can be simplified by directly using subsetting techniques without the need for the intermediate steps. Here's a simplified version of the code:

```{r}
library(palmerpenguins)

data("penguins")

adelie_penguins <- penguins[penguins$species == "Adelie", ]
```

In this simplified version:

1\. The \`penguins\$species == "Adelie"\` condition is used to directly subset the \`penguins\` dataset, selecting only the rows where the species is "Adelie".

2\. The filtered subset is stored in the variable \`adelie_penguins\`.

This code achieves the same result as the original code but is more concise and eliminates the need for the intermediate steps and the \`step_one\` function.

</br>

3.  Is the AI model's response to Question 2 in the {tidyverse} style (e.g., does the response use any {tidyverse} functions, the `%>%` pipe, etc.)?

    The code is not in the tidyverse style.

</br>

4.  If the AI model's response did not use the {tidyverse}, ask it to convert the code into {tidyverese} style. If your response already used the {tidyverse}, make a guess as to why it might have already done so. *\<--- just a thought experiment, don't worry too much about how you answer this!*

**Question to chat bot** (*if your previous question was not in tidyverse)*:

"Can you convert this to tidyverse?"

**Chat bot response:**

Certainly! In the tidyverse, you can use the \`dplyr\` package to achieve the same result with a more readable and expressive syntax. Here's how you can filter the \`penguins\` dataset to include only the rows where the species is "Adelie" using the tidyverse approach:

```{r}
library(tidyverse)

library(palmerpenguins)

data("penguins")

adelie_penguins <- penguins %>%

  filter(species == "Adelie")
```

In this code:

1\. \`penguins %\>%\` takes the \`penguins\` dataset as input and pipes it into the next operation.

2\. \`filter(species == "Adelie")\` filters the dataset to include only the rows where the \`species\` column is "Adelie".

3\. The result is stored in the variable \`adelie_penguins\`.

The tidyverse approach using the \`filter\` function from the \`dplyr\` package provides a more intuitive way to express data manipulation operations.

</br>

5.  Does the code that ChatGPT put together for you work? If it doesn't, ask the chat bot to fix it. Does it work now?

    ```{r}
    library(tidyverse)
    library(palmerpenguins)

    data("penguins")

    adelie_penguins <- penguins %>%
      filter(species == "Adelie")
    ```

    The tidyverse code chunk provided by ChatGPT works!

    </br>

6.  Ask the chat bot if there is a significant difference in bill length across Adelie penguin sexes in the {palmerpenguins} data set (using R).

**Question posed to chat bot:**

"Is there a significant difference in bill length across Adelie penguin sexes in the {palmerpenguins} data set (using R)?"

**Chat bot response:**

To test if there is a significant difference in bill length across Adelie penguin sexes in the \`palmerpenguins\` dataset, you can use a t-test. In R, you can perform an independent two-sample t-test using the \`t.test()\` function. Here's how you can do it for the Adelie penguins:

```{r}
# Load required library and data

library(palmerpenguins)

# Load data

data("penguins")

# Subset the data to include only Adelie penguins

adelie_penguins <- subset(penguins, species == "Adelie")

# Perform independent two-sample t-test for bill length across sexes

t_test_result <- t.test(bill_length_mm ~ sex, data = adelie_penguins)

# Print the t-test result

print(t_test_result)
```

In this code:

1\. The \`subset()\` function is used to filter the \`penguins\` dataset and include only the Adelie penguins.

2\. The \`t.test()\` function is applied to test if there is a significant difference in \`bill_length_mm\` across the \`sex\` variable (which represents the sexes of the Adelie penguins) within the \`adelie_penguins\` dataset.

3\. The t-test result, including the t-statistic, degrees of freedom, p-value, and confidence interval, is printed to the console.

The \`t.test()\` function will output the results of the t-test, including whether the difference in means is statistically significant. The p-value will indicate whether there is a significant difference in bill length across Adelie penguin sexes. If the p-value is below a chosen significance level (commonly 0.05), you can conclude that there is a significant difference in bill length between the sexes.

</br>

7.  Does the code produced work? (*Copy and paste the code in R Studio console to run it).* If not, what was wrong with it's answer?

    ```{r}
    # Load required library and data
    library(palmerpenguins)

    # Load data
    data("penguins")

    # Subset the data to include only Adelie penguins
    adelie_penguins <- subset(penguins, species == "Adelie")

    # Perform independent two-sample t-test for bill length across sexes
    t_test_result <- t.test(bill_length_mm ~ sex, data = adelie_penguins)

    # Print the t-test result
    print(t_test_result)
    ```

    The code seems to work!

    </br>

8.  Did the chat bot's response include an assessment of whether the Adelie penguins data met the assumptions of the statistical test used in the response? (e.g., test for normality or equal variances)? If not, pose the question again to test for the assumptions and paste the new question/response below.

    The chat bot's response didn't include an assessment of whether the data met the assumptions of the statistical test in the response, but chose a Welch test over a pooled t-test which can be used with unequal variance.

**Question to chat bot:**

"Does the data meet the assumptions for the statistical test used?"

**Chat bot response:**

To perform a t-test, certain assumptions need to be met. Here are the assumptions for an independent two-sample t-test:

1\. \*\*Normality:\*\* The data within each group (in this case, male and female Adelie penguins) should be approximately normally distributed. This assumption is more critical with smaller sample sizes. You can check for normality visually using histograms or quantile-quantile (Q-Q) plots, or use statistical tests like the Shapiro-Wilk test or Kolmogorov-Smirnov test.

2\. \*\*Homogeneity of Variance:\*\* The variances of the two groups (male and female Adelie penguins) should be equal. This assumption is important for the validity of the t-test. You can check for homogeneity of variance using statistical tests like Levene's test or visual methods like boxplots.

Before interpreting the results of the t-test, it's essential to assess whether these assumptions are met. You can use the following R code to check the assumptions for the t-test:

```{r}
# Check normality assumptions

shapiro.test(adelie_penguins$bill_length_mm[adelie_penguins$sex == "Male"])

shapiro.test(adelie_penguins$bill_length_mm[adelie_penguins$sex == "Female"])

# Check homogeneity of variance

leveneTest(bill_length_mm ~ sex, data = adelie_penguins)
```

In this code:

\- \`shapiro.test()\` is used to test the normality of bill lengths separately for male and female Adelie penguins.

\- \`leveneTest()\` from the \`car\` package is used to test the homogeneity of variance between male and female Adelie penguins.

If the normality tests (Shapiro-Wilk tests) indicate that the data within each group is approximately normally distributed (p \> 0.05), and the Levene's test suggests that the variances are not significantly different between the groups (p \> 0.05), then the assumptions for the t-test are reasonably met, and you can interpret the t-test results with confidence.
