---
title: "Hurricanes_10272023"
author: "Billy McLean"
date: "2023-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Packages

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)

```

Read in data

```{r}
data <- read_csv("C:/Users/Admin/Desktop/ESS523A/Week-2-Debugging-AI/data/storms.csv")
```

```{r}
glimpse(data)
```

Have storms increased in frequency?

```{r}
data %>% ggplot()+
  geom_histogram(mapping = aes(x = year), binwidth = .75)
```

Looks like it

How about windspeeds? They look to correlate with latitude

```{r}
data %>% ggplot()+
  geom_hex(mapping = aes(x = lat, y = wind))
```

It looks like most low and high windspeed evens occur 10 and 30 degrees latitude


```{r}
data %>% ggplot()+
  geom_hex(mapping = aes(x = long, y = wind))
```

This shows us the data is from the western hemisphere, not super useful

```{r}
status_year <- data %>% group_by(status, year) %>% 
  count()
```


How about Cat 5 storms specifically? Have we had more of those in recent years?
Filter the dataset to just cat 5 storms


```{r}
catfive <- data %>% filter(category == "5")
  
```

HISTOGRAM TIME

```{r}
catfive %>%  ggplot()+
  geom_histogram(mapping = aes(x = year))
```

Looks like more cat 5 storms since 2000, yikes

BUT
Cat 5 is a rare and horrible thing, lets give some credence to cat 3 and 4 storms as well

Do the same as above but for category 3, 4, and 5 storms

```{r}
highintensity <- data %>% filter(category == c(3, 4, 5))
  
```

HISTOGRAM TIME

```{r}
highintensity %>%  ggplot()+
  geom_histogram(mapping = aes(x = year))
```

SAME RESULT LETS ROCK AND ROLL

But am I smart enough to do anything with this knowledge? Probably not

Lets do the opposite and compare


```{r}
lowintensity <- data %>% filter(category == c(1, 2))
  
```

HISTOGRAM TIME

```{r}
lowintensity %>%  ggplot()+
  geom_histogram(mapping = aes(x = year))
```

At a glance it looks like there have been more low intensity storms as well. And a cat 2 still aint nothing to mess with really.

Lets break these out by category and see if it looks like things are increasing
```{r}
data %>%  ggplot()+
  geom_histogram(mapping = aes(x = year))+
  facet_wrap(~category, scales = "free")
  
```
Sure does, but storms can get upgraded as they progress, I could clean this data up to get better ideas but have run out of time. 


Group by year and category
```{r}
cat_year <- data %>% group_by(category, year) %>% 
  count()
```

filter out NA values
```{r}

filtered_catyear <- cat_year %>% na.omit()
  
```


This didnt really go anywhere in the end. We'll get em next time




