---
title: "How have storms changed?"
author: "Caitlin Mothes, PhD"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set up
```{r}
library(tidyverse)
library(sf) # spatial packages to make a map of storms
library(mapview)

```

Read in data
```{r}
storms <- read_csv("data/storms.csv")
```


Initial exploration
```{r}
glimpse(storms)

View(storms)
```


Make a map of storms
```{r}
storms_sf <- st_as_sf(storms, coords = c("long", "lat")) %>% 
  st_set_crs(4296)

mapview(storms_sf)
```


Lets look at wind and pressure over time

Each storm has values every 6 hours, summarize and find the max wind speed and pressure for each storm

But first, make sure there are not any duplicate storm names
```{r}
storms %>% 
  distinct(name, year) %>% 
  group_by(name) %>% 
  count() %>% 
  #filter if any count() is > 1, meaning duplicates
  filter(n > 1)
```

Back up, lets just look at those that made it to hurricane status
```{r}
hurricanes <- storms %>% 
  filter(status == "hurricane")
```

Check for duplicate named hurricanes again
```{r}
hurricanes %>% 
  # get distinct name/year
  distinct(name, year) %>% 
  # see how many dup names
  group_by(name) %>% 
  count() %>% 
  filter(n > 1) %>% 
  #sort to see which names reoccur the most
  arrange(-n)

# Earl!!
```

Now back to the wind and pressure values, find the max for each hurricane/year 
```{r}
storm_stats <- hurricanes %>%
  group_by(name, year) %>% 
  summarize(max_wind = max(wind), max_pressure = max(pressure))

# find storms with max recorded wind and pressure
storm_stats %>% 
  ungroup() %>% 
  filter(max_wind == max(max_wind) | max_pressure == max(max_pressure))
  
```


OK, back to change over time. 

How many hurricanes per year?
```{r}
hurricanes %>% 
  group_by(year) %>% 
  count() %>% 
  ggplot(aes(x = year, y = n)) +
  geom_line()
```

Max hurricane wind speed/year?
```{r}
hurricanes %>% 
  group_by(year) %>% 
  summarize(max_wind = max(wind)) %>% 
  ggplot(aes(x = year, y = max_wind)) +
  geom_line()
```

Average wind speed/year? (Using all values)
```{r}
hurricanes %>% 
  group_by(year) %>% 
  summarize(avg_wind = mean(wind)) %>% 
  ggplot(aes(x = year, y = avg_wind)) +
  geom_line()

#average is very different trend than looking at just the max!
```


