---
title: "storm_explorer"
author: "Katie Willi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

How have storms changed over time?

I have no idea! I'm going to start off by loading in the dataset, and creating a new column for the date:

```{r}
library(tidyverse)
library(sf)

storms <- read_csv('data/storms.csv') %>%
  mutate(date = lubridate::ymd(paste0(year,"-",month,"-",day))) %>%
  # in case a storm lasts thru the new year - group it with previous year
  # no idea how often this actually happes though
  mutate(storm_year = ifelse(month %in% c(1), year - 1, year))
```

Let's also visualize the storm tracks for fun:

```{r}
storms_sf <- storms %>%
  st_as_sf(coords = c("long", "lat"), crs = 4269)

mapview::mapview(storms_sf, zcol = "storm_year")
```

How many days do the hurricanes last? Is it changing?

```{r}
days <- storms %>%
  filter(status == "hurricane") %>%
  # only a single observation for every storm-date combo
  distinct(name, storm_year, date) %>%
  # names have been reused so must also group by year:
  group_by(name, storm_year) %>%
  # total days hurrican lasted
  summarize(count = n())

ggplot(days)+
  geom_point(aes(x=storm_year, y=count, color = name)) +
  theme_bw() +
  theme(legend.position = "none")
```

Get the max and min of some of our variables by year, and then plot them. This is something ChatGPT told me to do!

```{r}
storms_filtered <- storms %>%
  filter(status == "hurricane") %>% 
  select(name, storm_year, month, day, wind, pressure) %>%
  # remove instances where year is na
  filter(!is.na(storm_year))

# Calculate annual summaries
storms_summary <- storms_filtered %>%
  group_by(storm_year) %>%
  summarize(mean_wind = mean(wind, na.rm = TRUE),
            max_wind = max(wind, na.rm = TRUE),
            min_pressure = min(pressure, na.rm = TRUE),
            max_pressure = max(pressure, na.rm = TRUE))
# Create time series plots
ggplot(storms_summary, aes(x = storm_year)) +
  # Not a huge fan of developing plots this way. I prefer long data sets.
  geom_line(aes(y = mean_wind, color = "Mean Wind Speed")) +
  geom_line(aes(y = max_wind, color = "Max Wind Speed")) +
  geom_line(aes(y = min_pressure, color = "Min Pressure")) +
  geom_line(aes(y = max_pressure, color = "Max Pressure")) +
  labs(title = "Hurricane Trends Over Time",
       x = "Year",
       y = "Values",
       color = "Variables") +
  scale_color_manual(values = c("Mean Wind Speed" = "pink", 
                                "Max Wind Speed" = "red", 
                                "Min Pressure" = "maroon", 
                                "Max Pressure" = "darkred")) +
  theme_bw()
```

Have certain types of storms become more frequent with time?

```{r}
storm_dist <- storms %>% 
  # create a single observation for a given storm, its storm type (`status`), and day
  distinct(name, status, storm_year, date) 

# this plot shows the frequency (in days) of a given storm type by date
ggplot(data = storm_dist) +
  geom_histogram(aes(x = date), color = "darkred",fill = "maroon") +
  facet_wrap(~status) +
  theme_bw()
```

# Are storms increasing in the North Atlantic?

```{r}
upper <- storms %>%
  #filter(status == "hurricane") %>% 
  filter(lat >= 42) %>%
  distinct(name, storm_year) %>%
  mutate(combo = paste0(name,storm_year))


upper_trails <- storms %>%
  #filter(status == "hurricane") %>% 
  mutate(combo = paste0(name,storm_year)) %>%
  filter(combo %in% upper$combo) %>%
  st_as_sf(coords = c("long","lat"), crs = 4269)
  
mapview(upper_trails, zcol = "name")

ggplot(data = upper_trails) +
  geom_histogram(aes(x = date), color = "darkred",fill = "maroon") +
  facet_wrap(~status) +
  theme_bw()
```

Are they moving further north?
```{r}
upper <- storms %>%
  filter(status == "hurricane") %>% 
  group_by(storm_year) %>%
  summarize(max_lat = max(lat))

ggplot(data = upper) +
  geom_point(aes(x=storm_year, y= max_lat)) +
  geom_smooth(aes(x=storm_year, y = max_lat), method = "lm") +
  theme_bw()
```

# Are storms reaching inland more frequently?

```{r}
west <- storms %>%
  #filter(status == "hurricane") %>% 
  filter(long <=  -81) %>%
  distinct(name, storm_year) %>%
  mutate(combo = paste0(name,storm_year))

west_trails <- storms %>%
  #filter(status == "hurricane") %>% 
  mutate(combo = paste0(name,storm_year)) %>%
  filter(combo %in% west$combo) %>%
  st_as_sf(coords = c("long","lat"), crs = 4269)
  
mapview(west_trails, zcol = "storm_year")

ggplot(data = west_trails) +
  geom_histogram(aes(x = date), color = "darkred",fill = "maroon") +
  facet_wrap(~status) +
  theme_bw()
```

Are they generally moving further west?
```{r}
west <- storms %>%
  #filter(status == "hurricane") %>% 
  group_by(storm_year) %>%
  summarize(max_long = max(long))

ggplot(data = west) +
  geom_point(aes(x=storm_year, y= max_long)) +
  geom_smooth(aes(x=storm_year, y = max_long), method = "lm") +
  theme_bw()
```
