---
title: "Week-2-notes"
author: "Erika Lee"
date: "2023-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
source("setup.R")
```

## Week 2 Notes

### How to fix errors:

-   How to see which type of data is in a column: glimpse(dataname) - tidyverse version of viewing types of data

-   How to create a new column: Example

    -   cpf \<- cpr %\>%

        mutate(year = lubridate::year(Date) \> "2022") - mutating then filtering our the years that are greater than 2022...

But first, have to make Date a year...

-   mutate (Date = lubridate::dmy(Date))

-   Therefore:

    mutate(Date = lubridate::dmy(Date)) %\>%

    mutate(year = lubridate::year(Date)) %\>%

    filter(year \> "2022") %\>%

    filter(grepl("Poudre", site))

    \*\*grepl() - is a way of filtering a word within a column (site is the column, Poudre is the word that exists within the column)

-   Package load is reloading each package every time you run the code, therefore... make sure to comment out the package.Load code line after it has been run.

-   Make sure code chunks are closed, and are in fact IN a code chunk. And, don't have nested code chunks

-   R doesn't like spaces... keep a system for lowercase and use either an \_ or . or -, and be consistent with this!
