---
title: "Code Review Desired Output"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

In this workflow, we want to explore the Palmer Penguins data set collected at the Palmer Station LTER in Antarctica.

```{r}
library(tidyverse)
library(palmerpenguins)
```

We read in the data and initially explore its structure:

```{r}
data("penguins")
summary(penguins)
```

Lets first explore body mass among all the penguin species with a histogram:

```{r warning=FALSE}
penguins %>% 
  ggplot()+
  geom_histogram(aes(x = body_mass_g))
```

Now we want to visualize the relationship between body mass and bill length *among species*:

```{r warning=FALSE}
penguins %>%
  ggplot() +
  geom_point(aes(x = body_mass_g, y = bill_length_mm, color = species))
```

Finally, lets create a new data frame called `adelie` that contains data for just the *Adelie* penguin species, and then visualize the differences in that species' body mass among islands.

```{r}
adelie <- penguins %>% 
  filter(species == 'Adelie')
```

```{r warning=FALSE}
adelie %>% 
  ggplot() +
  geom_boxplot(aes(x = island, y = body_mass_g))
```

This figure suggests *Adelie* individuals on the Dream island are slightly smaller that those on the Biscoe and Torgersen islands.
