---
title: "Code Review/Debugging Assignment"
author: 
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## The Assignment:

*Tip: View this in the visual editor in RStudio*

You want to re-create the HTML document in the `lessons/` folder in this repo titled `code_debugging_output.html`. Below is the workflow to do so, ***however*** you will notice if you attempt to render this R Markdown document as is it does NOT reproduce the `code_debugging_output.html` file, and won't even render until you fix some of the mistakes/missing elements. Therefore, this workflow is not reproducible! Your task is to fix the below coding workflow so that it renders and **exactly** matches the output of `code_debugging_output.html` (*including* whether or not certain code chunks and/or outputs are shown).

There are a total of **5 fixes** you need to make to the workflow below. To complete this assignment you must ***1)*** Identify the 5 errors and specify what they are below (*4 pts each*) and ***2)*** fix the errors and render this entire .Rmd as am HTML doc where the content below matches that of the `code_debugging_output.html` (*5 pts*).

**List here the 5 errors you had to fix in this workflow:**

**1. Library Lapse**:

-   I skipped the opening act! Forgot to load the showstopper libraries like **`tidyverse`** and **`palmerpenguins`**. Without them, it's like trying to bake without flour.

**2. Name Game Goof**:

-   Oops! Called our penguin friend by the wrong name. Instead of waving to 'Adelei', I should've hailed 'Adelie'. Names matter, especially at a penguin party!

**3. `adelie` Amnesia**:

-   In my excitement to filter data, storing it was overlooked in the **`adelie`** dataframe. It's like making tea and then forgetting to drink it.

**4. Histogram Hesitation**:

-   The histogram was acting up because I didn't specify the **`binwidth`** or **`bins`**. It's like forgetting the rhythm of a dance.

**5. Point Plot's Missing Palette**:

-   I was scratching my head on this one. When visualizing the relationship between body mass and bill length, the addition of **`color`** aesthetic was overlooked to differentiate the species. It's like missing the sprinkles on a sundae!

With these fixes, the workflow should be smoother than a penguin's slide on ice!

## The workflow you need to fix begins here:

In this workflow, we want to explore the Palmer Penguins data set collected at the Palmer Station LTER in Antarctica. This data is stored in the {palmerpenguins} data package.

We read in the data and initially explore its structure:

```{r}

# Error 1: Loading the necessary libraries
library(tidyverse)
library(palmerpenguins)

```

```{r}
data("penguins")
summary(penguins)
```

Let's first explore body mass among all the penguin species with a histogram:

```{r}
# Error 4: Specifying binwidth for histogram
penguins %>% 
  ggplot() +
  geom_histogram(aes(x = body_mass_g), binwidth = 500)

#old
#{r warning=FALSE}
#penguins %>% 
#  ggplot()+
#  geom_histogram(aes(x = body_mass_g))

```

Now we want to visualize the relationship between body mass and bill length *among species*:

```{r}

# Error 5: Adding color aesthetic to differentiate species
penguins %>%
  ggplot() +
  geom_point(aes(x = body_mass_g, y = bill_length_mm, color = species))

#old
#{r warning=FALSE}
#penguins %>%
#  ggplot() +
#  geom_point(aes(x = body_mass_g, y = bill_length_mm))

```

Finally, let's create a new data frame called `adelie` that contains data for just the *Adelie* penguin species, and then visualize the differences in that species' body mass among islands.

```{r}

# Errors 2 & 3: Fixing the species name and assigning the filtered data to adelie dataframe
adelie <- penguins %>% 
  filter(species == 'Adelie')

#old
#penguins %>% 
#  filter(species == 'Adelei')

```

```{r}

adelie %>% 
  ggplot() +
  geom_boxplot(aes(x = island, y = body_mass_g))

#old
#{r warning=FALSE}
#adelie %>% 
#  ggplot() +
#  geom_boxplot(aes(x = island, y = body_mass_g))

```

This figure suggests *Adelie* individuals on the Dream island are slightly smaller that those on the Biscoe and Torgersen islands.
