---
title: "Code Review/Debugging Assignment"
author: 
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## The Assignment:

*Tip: View this in the visual editor in RStudio*

You want to re-create the HTML document in the `lessons/` folder in this repo titled `code_debugging_output.html`. Below is the workflow to do so, ***however*** you will notice if you attempt to render this R Markdown document as is it does NOT reproduce the `code_debugging_output.html` file, and won't even render until you fix some of the mistakes/missing elements. Therefore, this workflow is not reproducible! Your task is to fix the below coding workflow so that it renders and **exactly** matches the output of `code_debugging_output.html` (*including* whether or not certain code chunks and/or outputs are shown).

There are a total of **5 fixes** you need to make to the workflow below. To complete this assignment you must ***1)*** Identify the 5 errors and specify what they are below (*4 pts each*) and ***2)*** fix the errors and render this entire .Rmd as am HTML doc where the content below matches that of the `code_debugging_output.html` (*5 pts*).

**List here the 5 errors you had to fix in this workflow:**

1.  The palmer penguin package was not initially loaded in, as well as the tidyverse package. 

2. Made summary section visible, but added {r warning=FALSE} so that these results don't show up towards the end.

3.  Colors were not assigned in geom_point (Fixed it to be: (aes(x = body_mass_g, y = bill_length_mm, color = species))

4.  It should be 'Adelie' instead of 'Adelei' in the code.

5.  The last code chunk requires a data frame prior to running the code. (adding adelie \<- to the code should do.)

## The workflow you need to fix begins here:

In this workflow, we want to explore the Palmer Penguins data set collected at the Palmer Station LTER in Antarctica. This data is stored in the {palmerpenguins} data package.

We read in the data and initially explore its structure:

```{r}
library(palmerpenguins)
library(tidyverse)
data("penguins")
```

```{r warning=FALSE}
summary(penguins)
```

Let's first explore body mass among all the penguin species with a histogram:

```{r warning=FALSE}
penguins %>% 
  ggplot()+
  geom_histogram(aes(x = body_mass_g))
```

Now we want to visualize the relationship between body mass and bill length *among species*:

```{r warning=FALSE}
penguins %>%
  ggplot() +
  geom_point(aes(x = body_mass_g, y = bill_length_mm, color = species))
```
Finally, let's create a new data frame called `adelie` that contains data for just the *Adelie* penguin species, and then visualize the differences in that species' body mass among islands.

```{r warning=FALSE}
adelie <- penguins %>% 
  filter(species == 'Adelie')
```



```{r warning=FALSE}
adelie %>% 
  ggplot() +
  geom_boxplot(aes(x = island, y = body_mass_g))
```

This figure suggests *Adelie* individuals on the Dream island are slightly smaller that those on the Biscoe and Torgersen islands.
