---
title: "Code Review/Debugging Assignment"
author: 
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## The Assignment:

*Tip: View this in the visual editor in RStudio*

You want to re-create the HTML document in the `lessons/` folder in this repo titled `code_debugging_output.html`. Below is the workflow to do so, ***however*** you will notice if you attempt to render this R Markdown document as is it does NOT reproduce the `code_debugging_output.html` file, and won't even render until you fix some of the mistakes/missing elements. Therefore, this workflow is not reproducible! Your task is to fix the below coding workflow so that it renders and **exactly** matches the output of `code_debugging_output.html` (*including* whether or not certain code chunks and/or outputs are shown).

There are a total of **5 fixes** you need to make to the workflow below. To complete this assignment you must ***1)*** Identify the 5 errors and specify what they are below (*4 pts each*) and ***2)*** fix the errors and render this entire .Rmd as am HTML doc where the content below matches that of the `code_debugging_output.html` (*5 pts*).

**List here the 5 errors you had to fix in this workflow:**

1.  Called palmerpengiuns package with library(). 

2.  Read in dplyr and ggplot2 packages to use the pipe (%>%) and ggplot(). 

3.  Added "color = species" after x and y variables in the aes() function within ggplot. This defined the dots for each species by color and added a legend for reference. 

4.  Assigned the new data frame to the "adelie" name using <- before the rest of the code. 

5.  Fixed the typo in the species name to "Adelie" as it is within the data when creating the new dataframe within the filter() function. 

## The workflow you need to fix begins here:

In this workflow, we want to explore the Palmer Penguins data set collected at the Palmer Station LTER in Antarctica. This data is stored in the {palmerpenguins} data package.

We read in the data and initially explore its structure:

```{r}
library(palmerpenguins)
library(dplyr)
data("penguins")
summary(penguins)
```

Let's first explore body mass among all the penguin species with a histogram:

```{r warning=FALSE}
library(dplyr)
library(ggplot2)

penguins %>% 
  ggplot()+
  geom_histogram(aes(x = body_mass_g))
```

Now we want to visualize the relationship between body mass and bill length *among species*:

```{r warning=FALSE}
penguins %>%
  ggplot() +
  geom_point(aes(x = body_mass_g, y = bill_length_mm, color = species))
```

Finally, let's create a new data frame called `adelie` that contains data for just the *Adelie* penguin species, and then visualize the differences in that species' body mass among islands.

```{r}
adelie <- penguins %>% 
  filter(species == 'Adelie')
```

```{r warning=FALSE}
adelie %>% 
  ggplot() +
  geom_boxplot(aes(x = island, y = body_mass_g))
```

This figure suggests *Adelie* individuals on the Dream island are slightly smaller that those on the Biscoe and Torgersen islands.
